Functional oxide layers for electrical isolation and chemical passivation of steel substrates

### 2023-02-12 Thesis/chapter/031-theory-calculations.tex
nearly finish theory

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Why introduce machine learning to this project? 
%I wanted to apply and delve into what I've been studying during my undergrad courses. 
%This seemed like the perfect oppurtunity. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Is machine learning just hyped statistics? No! 
		%\href
		%{https://doi.org/10.1177\%2F1352458520978648}
		%{Machine and deep learning in MS research are just powerful statistics – No}
What is \gls{ml}? Some might argue that \gls{ml} is just overglorified statistics, but that's not the case. 
Although \gls{ml} uses several statistical methods, their goal and frame conditions are different. 
%\td{The goal of machine learning is to ???}
%
%\begin{itemize}
%	\item \td{let GPT-3 write}
%	\item check out links at \url{https://yewtu.be/watch?v=PqbB07n\_uQ4}
%	\item see pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
%	\item 
%		\href
%		{https://doi.org/10.1177\%2F1352458520978648}
%		{Machine and deep learning in MS research are just powerful statistics – No}
%\end{itemize}
%
%\td{
%\url{https://stats.stackexchange.com/questions/5026/what-is-the-difference-between-data-mining-statistics-machine-learning-and-ai}
%In principle both \gls{ml} and statistics use math to get information out of data. 
%AI might use machine learning to create "intelligent" acting (playing a game or driving a  car). 
\Gls{ml} tries to predict unseen data points and statistics is a subfield of maths 
which tries to get insight into a given data. 
For example, in a statistical model, it is desirable to reduce the number of inputs. 
This allows the statistician to better study how a change in the input variable can be 
directly affected by the output variable.\cite{gontcharov2019}
%pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
%}
%What is statistics, though, and what seperates it from \gls{ml}? 
More precisely, mathematical statistics (stochastic)\cite{haertler2014statistisch} is the subject of of 
finding mathematical models to describe the data 
whereas (classical) statistics is the domain of representing the data. 
\td{talk about statistics, then the difference to Ml} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\td{if i want to learn about support vector machine, read chapter 4,7,8 from "Learning from Data" by Cherkassky\cite{cherkassky1998learning} (TUBib 726-1998)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
%
%\todo{check out SVM and KRR}
%
\Gls{ml} is at the base of most \gls{ai}s.
It is an umbrella term for programs with instructions to learn from data, i.e. gain knowledge, categorise, predict and make decisions based on data. 
%
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous data) versus classification (discrete, categorical data). 
%Two architecture which don't quite fit into this classification are Generative Adversarial Network (GAN) and jk
Independent from these $2\times2$ categories there are multiple ways to let machines learn from data.
%
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and are useful in quantum mechanics too\cite{westermayr2020combining}.
Other common architectures include linear regression, kernel ridge regression and support vector regression.
There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
\gls{pso} algorithms take advantage of the ability to cope with local optima by evolving several candidate solutions simulta- neously\cite{villanova2010function}.
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
%\td{This is exactly what I needed because every experiment is very time intensive.}

A genetic algorithm (GA) is a search algorithm that uses principles of natural selection and genetics to optimize a search space. A GA starts with a population of randomly generated solutions, or chromosomes, and then proceeds to breed them together to create new solutions. The new solutions are then tested for fitness, and the best solutions are selected to create the next generation of chromosomes. This process is repeated until a satisfactory solution is found.\footnote{This paragraph was written by GPT\cite{Liu2021gpt} given the input "Introduction to genetic algorithms: "}

@@

\Gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behaviour of bird flocks and fish schools\cite{Kennedy1995,villanova2010function}.
%"PSO is a population based heuristic inspired by the flocking behavior of birds. 
%To simulate the behavior of a swarm, each bird (or particle) is allowed to fly towards the optimum solution."\cite{villanova2010function}
Each particle has an associated position and velocity. 
Every movement across the search space is 
additionally to a stochastic term
%stochastic and 
influenced by its particle velocity and position as well as its and the swarm's best visited position.

@@

\subsubsection{Kernel Ridge Regression and Support Vector Machine}

\textbf{goal}
\begin{itemize}
    \item intro to stat and ml 
    \item say that although different will be abgehandelt in the same chapter 
    \item basic statistics: MAE vs MSE, R, std, F-test 
    \item lin reg 
    \item ridge regression 
    \item kernel ridge regression 
    \item support vector machine 
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but don't tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimisation, actually incorporate prediction into selection process but don't per se specify which pred and optimisation)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}

\textbf{current}
\begin{itemize}
    \item ML vs statistics
    \item AI vs ML 
    \item ML methods
    \item EMMA 
    \item KRR 
    \item DOE
    \item ANOVA
\end{itemize}

@@

%%% FEATURE SELECTION 
% https://stats.stackexchange.com/questions/39243/how-does-one-interpret-svm-feature-weights
% gene (feature) selection with SVM https://link.springer.com/content/pdf/10.1023/A:1012487302797.pdf
% Intro to feature selection https://axon.cs.byu.edu/Dan/778/papers/Feature%20Selection/guyon2.pdf

\td{statistical methods to bewerten a model are: cross validation, leave-one-out validation ( special case of cross validation), correlation coefficient for linear fits, mean square error. methods to fit data are: lin reg, polynomial regression, ridge regression, KRR, SVM and?}

>>>>

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Why introduce machine learning to this project? 
%I wanted to apply and delve into what I've been studying during my undergrad courses. 
%This seemed like the perfect oppurtunity. 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%Is machine learning just hyped statistics? No! 
		%\href
		%{https://doi.org/10.1177\%2F1352458520978648}
		%{Machine and deep learning in MS research are just powerful statistics – No}
What is \gls{ml}? Some might argue that \gls{ml} is just overglorified statistics, but that's not the case. 
Although \gls{ml} uses several statistical methods, their goal and frame conditions are different. 
%\td{The goal of machine learning is to ???}
%
%\begin{itemize}
%	\item \td{let GPT-3 write}
%	\item check out links at \url{https://yewtu.be/watch?v=PqbB07n\_uQ4}
%	\item see pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
%	\item 
%		\href
%		{https://doi.org/10.1177\%2F1352458520978648}
%		{Machine and deep learning in MS research are just powerful statistics – No}
%\end{itemize}
%
%\td{
%\url{https://stats.stackexchange.com/questions/5026/what-is-the-difference-between-data-mining-statistics-machine-learning-and-ai}
%In principle both \gls{ml} and statistics use math to get information out of data. 
%AI might use machine learning to create "intelligent" acting (playing a game or driving a  car). 
\Gls{ml} tries to predict unseen data points and statistics is a subfield of maths 
which tries to get insight into a given data. 
For example, in a statistical model, it is desirable to reduce the number of inputs. 
This allows the statistician to better study how a change in the input variable can be 
directly affected by the output variable.\cite{gontcharov2019}
%pic from \url{https://towardsdatascience.com/notes-on-artificial-intelligence-ai-machine-learning-ml-and-deep-learning-dl-for-56e51a2071c2}
%}
%What is statistics, though, and what seperates it from \gls{ml}? 
More precisely, mathematical statistics (stochastic)\cite{haertler2014statistisch} is the subject of 
finding mathematical models to describe the data 
whereas (classical) statistics is the domain of representing the data. 
\td{talk about statistics, then the difference to Ml} 
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
%\td{if i want to learn about support vector machine, read chapter 4,7,8 from "Learning from Data" by Cherkassky\cite{cherkassky1998learning} (TUBib 726-1998)}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

@@

%
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsubsection{Machine Learning Methods}
%
%\todo{check out SVM and KRR}
%
\Gls{ml} is at the base of most \gls{ai}s.
It is an umbrella term for programs with instructions to learn from data, i.e. gain knowledge, categorise, predict and make decisions based on data. 
%
There is a platitude of different machine learning methods: 
they can be divided into supervised (training set is labeled) and unsupervised (exploratory).  
An orthogonal division can be made by regression (continuous data) versus classification (discrete, categorical data). 
%Two architecture which don't quite fit into this classification are Generative Adversarial Network (GAN) and jk
Independent from these $2\times2$ categories there are multiple ways to let machines learn from data.
%
\Gls{nn} (one of the most popular architectures for big data\cite{Chiroma2019}) are loosely modeled after the brain\cite{bishop1994neural}.
Artificial neurons (also called nodes), which are arranged in layers, 
are connected to each of the neurons of previous and next layers
and the weights (parameters of intensity), with which the data is routed from one neuron to another, 
are optimised during training. 
%
Convolutional-layer \gls{nn} excel in picture recognition\cite{Lecun1995conv} and are useful in quantum mechanics too\cite{westermayr2020combining}.
Other common architectures include linear regression, kernel ridge regression and support vector regression.

\input{chapter/033-linreg-krr-svm.tex}

There are also lesser know algorithms like 
evolutionary algorithms (e.g. \gls{ga} and \gls{pso}).
\gls{pso} algorithms take advantage of the ability to cope with local optima by evolving several candidate solutions simultaneously\cite{villanova2010function}.
%
One advantage of evolutionary algorithms is that they start with a small data set
and periodically request new data in order to solve the problem iteratively.
%\td{This is exactly what I needed because every experiment is very time intensive.}

A \gls{ga} is a search algorithm that uses principles of natural selection and genetics to optimize a search space. A GA starts with a population of randomly generated solutions, or chromosomes, and then proceeds to breed them together to create new solutions. The new solutions are then tested for fitness, and the best solutions are selected to create the next generation of chromosomes. This process is repeated until a satisfactory solution is found.\footnote{This paragraph was written by GPT\cite{Liu2021gpt} given the input "Introduction to genetic algorithms: "}

@@

A \gls{pso} also uses a starting population of particles where each experiment (particle) 
is represented by its independent and dependent variables. 
It was originally inspired by the behaviour of bird flocks and fish schools\cite{villanova2010function,Kennedy1995}.
%"PSO is a population based heuristic inspired by the flocking behavior of birds. 
%To simulate the behavior of a swarm, each bird (or particle) is allowed to fly towards the optimum solution."\cite{villanova2010function}
Each particle has an associated position and velocity. 
Every movement across the search space is, 
additionally to a stochastic term, 
%stochastic and 
influenced by its particle velocity and position as well as its and the swarm's best visited position.

@@

\iffalse
\subsubsection{Kernel Ridge Regression and Support Vector Machine}
\textbf{goal}
\begin{itemize}
    \item intro to stat and ml 
    \item say that although different will be abgehandelt in the same chapter 
    \item basic statistics: MAE vs MSE, R, std, F-test 
    \item lin reg 
    \item ridge regression 
    \item kernel ridge regression 
    \item support vector machine 
    \item anova (categorical), PCA, step wise regression
    \item all previous ananlyse data to predict, but don't tell me which data point is needed next 
    \item DOE (add data acquisition)
    \item genetic algorithm and pso (they combine data gen with regression/prediction and optimisation, actually incorporate prediction into selection process but don't per se specify which pred and optimisation)
    \item comment: always tradeoff between complexity and interpretability
    \item quesiton: which paragraph by GPT2?? 
\end{itemize}

\textbf{current}
\begin{itemize}
    \item ML vs statistics
    \item AI vs ML 
    \item ML methods
    \item EMMA 
    \item KRR 
    \item DOE
    \item ANOVA
\end{itemize}
\fi

@@

%%% FEATURE SELECTION 
% https://stats.stackexchange.com/questions/39243/how-does-one-interpret-svm-feature-weights
% gene (feature) selection with SVM https://link.springer.com/content/pdf/10.1023/A:1012487302797.pdf
% Intro to feature selection https://axon.cs.byu.edu/Dan/778/papers/Feature%20Selection/guyon2.pdf
\subsubsection{Validation}
Validation is an important part of data analysis as it allows to estimate model errors. 
Often used methods are cross validation, leave-
statistical methods to bewerten a model are: 
cross validation, leave-one-out validation ( special case of cross validation), correlation coefficient for linear fits, mean square error, mean absolute erro
% https://www.researchgate.net/profile/Daniel-Berrar/publication/324701535_Cross-Validation/links/5cb4209c92851c8d22ec4349/Cross-Validation.pdf
% https://www.sciencedirect.com/science/article/abs/pii/0304380093901052
% https://www.eng.auburn.edu/~aesmith/files/janchap.pdf
% https://www.sciencedirect.com/science/article/pii/S0379073806003239
\td{cite from Hawkins 2004\cite{hawkins2004overfitting}}

### 2023-02-12 Thesis/chapter/032-alt-theory-calculations.tex
nearly finish theory

Thus maximizing the margin is equivalent with minimizing $\|\mathbf{w}\|$. 

@@

which can be rewritten by some mathematical wizardry (setting the partial derivatives of the lagrangian 
$\frac{\partial \mathcal{L}}{\partial \mathbf{w}}$ and $\frac{\partial \mathcal{L}}{\partial b}$ to zero and inserting into eq. (\ref{eq:svm03})): 

@@

	\mathcal{L} = \sum_i^n \alpha_i - \frac{1}{2} \sum_i^n \sum_i^n \alpha_i \alpha_j y_i y_j \mathbf{x}_i \cdot \mathbf{x}_j 

>>>>

Thus maximizing the margin is equivalent with minimizing $\|\mathbf{w}\|$ and minimizing $\frac{1}{2}\|\mathbf{w}\|^2$.

@@

which can be rewritten by some mathematical wizardry (setting the partial derivatives of the Lagrangian function
$\frac{\partial \mathcal{L}}{\partial \mathbf{w}}$ and $\frac{\partial \mathcal{L}}{\partial b}$ 
to zero and inserting into eq. (\ref{eq:svm03}))\cite{winston1992artificial,cherkassky1998learning}: 

@@

	\mathcal{L} = \sum_i^n \alpha_i - \frac{1}{2} \sum_i^n \sum_j^n \alpha_i \alpha_j y_i y_j \mathbf{x}_i \cdot \mathbf{x}_j 
